BioTime

### 2019-07-01 data-raw/biotime.Rmd
modify to include all taxa in biotime data

data <- read_csv(system.file("extdata", "biotime/biotime_metadata.csv", package = "biodivTS")) %>% 
  filter(taxa %in% c("Birds", "Mammals")) %>%
  select(study_id, taxa) %>%
  left_join(read_csv(system.file("extdata", "biotime/biotime_query.csv", package = "biodivTS")), by = "study_id")
```

Check which authority would result in the most matches 
```{r}
##Counts show that OTT is the best authority for both data sets
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
```

@@

```{r}
#Getting matches by scientific name 
biotime <- data %>% 
  #filter(taxa == "Birds") %>%
  select(genus_species) %>%
  drop_na() %>%
  distinct() %>%
  mutate(id = get_ids(genus_species, "itis")) %>%
  left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}
unres <- get_dupe_ids(biotime_ids, "sort")
unres 
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join ID's back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID from the two Warbler options
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

@@

#match all the un-ID'd names to other providers, check if synonyms given by those providers match known ITIS species 
alt_names <- map_df(providers, function(name) synonyms(unmatched, name)) %>%
  drop_na(acceptedNameUsageID) %>%
  #new matches can be in either the synonym or acceptedNameUsage column, so let's combine them
  gather(type, altProviderName, synonym, acceptedNameUsage) %>%
  select(altProviderName, input) %>% 
  distinct()

#match accepted names from 
syn_res <- by_name(na.omit(unique(alt_names$altProviderName)),"itis") %>%
  drop_na(acceptedNameUsageID) %>% 
  rename(altProviderName = input) %>%
  left_join(alt_names, na_matches = "never")

#some names have multiple matches (same ID but multiple scinames), have to pick one
## in this case all the duplicates have an accepted and a synonym, but there are some single matches that are just synonyms
## so we have to find duplicate ID's and then filter
alt_dupes <- syn_res %>% 
  group_by(input) %>%
  filter(n() > 1)

#filter duplicates for accepted, and replace old entries 
syn_res <- alt_dupes %>% 
  filter(taxonRank == "species", taxonomicStatus == "accepted") %>% 
  bind_rows(syn_res %>% 
              filter(!acceptedNameUsageID %in% alt_dupes$acceptedNameUsageID)) %>%
#then finish clean up 
  select(id = acceptedNameUsageID, input, scientificName) %>% 
  distinct() %>%
  left_join(data %>% select(-species, -genus), by = c("input" = "genus_species"), na_matches =  "never") %>%
  rename(matchingName = input) %>%
  ungroup()

#Below add's more rows, should just replace them 
biotime_data <- biotime_data %>% filter(!matchingName %in% syn_res$matchingName) %>%
  bind_rows(syn_res)

>>>>

data <- read_csv(system.file("extdata", "biotime/biotime_metadata.csv", package = "biodivTS")) %>% 
  select(study_id, taxa) %>%
  left_join(read_csv(system.file("extdata", "biotime/biotime_query.csv", package = "biodivTS")), by = "study_id")
```

Check which authority would result in the most matches 
```{r}
##Counts show that OTT is the best authority 
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
```

@@

```{r}
# #Getting matches by scientific name 
# biotime <- data %>% 
#   select(genus_species) %>%
#   drop_na() %>%
#   distinct() %>%
#   mutate(id = get_ids(genus_species, "itis")) %>%
#   left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want.
```{r}
unres <- get_dupe_ids(biotime_ids, "sort")
unres 
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
  left_join(biotime_ids) %>% #join ID's back with all the columns
  filter(acceptedNameUsageID %in% c(
    "ITIS:159807", #Heterodontus zebra, all the studies including it are using sci names, so we should match to the sci name match not common name match
    "ITIS:175125" #The Redhead is coming from bird studies, so choose the bird match, 
    #the rest can't be match due to too little information
  )) %>%
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

@@

#match all the un-ID'd names to other providers, check if synonyms given by those providers match known ITIS species 
alt_names <- map_df(providers, function(name) synonyms(unmatched, name)) %>%
  drop_na(acceptedNameUsageID) %>%
  #new matches can be in either the synonym or acceptedNameUsage column, so let's combine them
  gather(type, altProviderName, synonym, acceptedNameUsage) %>%
  select(altProviderName, input) %>% 
  distinct() %>%
  drop_na(altProviderName)

#match accepted names from 
syn_res <- by_name(na.omit(unique(alt_names$altProviderName)),"itis") %>%
  drop_na(acceptedNameUsageID) %>% 
  rename(altProviderName = input) %>%
  left_join(alt_names, na_matches = "never", by = "altProviderName")

#some names have multiple matches (same ID but multiple scinames), have to pick one
## in this case all the duplicates have an accepted and a synonym, but there are some single matches that are just synonyms
## so we have to find duplicate ID's and then filter
alt_dupes <- syn_res %>%
  select(-taxonID) %>%
  filter(taxonRank == "species") %>%
  group_by(input) %>%
  filter(n() > 1)

# #filter duplicates for accepted, and replace old entries 
# syn_res <- alt_dupes %>% 
#   filter(acceptedNameUsageID %in% c(
#     ITIS:331277, 
#   )) %>% 
#   bind_rows(syn_res %>% 
#               filter(!acceptedNameUsageID %in% alt_dupes$acceptedNameUsageID)) %>%
# #then finish clean up 
#   select(id = acceptedNameUsageID, input, scientificName) %>% 
#   distinct() %>%
#   left_join(data %>% select(-species, -genus), by = c("input" = "genus_species"), na_matches =  "never") %>%
#   rename(matchingName = input) %>%
#   ungroup()

#Just get rid of those with duplicates, deal with them later
add_syn_res <- syn_res %>% 
  filter(!input %in% alt_dupes$input) %>% #excluding duplicate matched species
  select(id = acceptedNameUsageID, input, scientificName) %>%
  distinct() %>%
  left_join(data %>% select(-species, -genus), by = c("input" = "genus_species"), na_matches =  "never") %>%
  rename(matchingName = input) %>%
  ungroup()

biotime_data <- biotime_data %>% 
  filter(!matchingName %in% add_syn_res$matchingName) %>%
  distinct() %>%
  bind_rows(add_syn_res)
